What we {\sf eval} in the shadows

### 2021-04-01 paper/main.tex
more

This section explains our methodology for selecting the corpus that
will be analyzed and how we implemented our dynamic analysis.

@@

Our corpus is assembled out of two set of programs, one consisting of libraries
obtained from \emph{CRAN} (Comprehensive R Archive Network\footnote{CRAN is the
  largest repository of R code with almost \CranAvailablePackagesRnd packages.
  It receives about 6 new package submissions a day~\cite{Ligges2017}. \cf
  \url{https://cran.r-project.org/web/packages/index.html}}) and the other from
the \emph{Kaggle} data science competition website\footnote{\cf
  \url{https://kaggle.com}}. The intent here is to contrast code written by
experienced developers (CRAN) with code authored by typical end users of the
language (Kaggle). Another difference is that CRAN is a curated repository. Each
submitted package must abide to a number of well-formedness rules that are
automatically checked asserting certain quality. Most relevant for this work is
that all of the runnable code is tested and only a successfully running package
is admitted in the archive. Kaggle on the other hand is more like GitHub, anyone
can submit any code. All the code in our corpus is runnable, \ie we have scripts
that invoke the CRAN packages and the Kaggle programs have input data.

% Clearly code coverage will vary; as is the case for any dynamic analysis, the
% quality of results is predicated on how representative the executions we can
% observe are with respect to all possible runs of the corpus. To mitigate the
% threat to validity attached to limited coverage, we obtain clients that will
% generate multiple \emph{runs} of CRAN packages. In this context, a run is the
% execution of a CRAN package or Kaggle program with a particular set of input
% values.

% \begin{figure}[!h]\centering\includegraphics[width=.7\linewidth]
%   {corpus.pdf}\caption{CRAN packages}\label{fig:corpus}
% \end{figure}

\mypara{CRAN Packages.} We used \CranAllPackages packages. These are almost all
of the available packages except for a few that we could not install due to
missing native dependencies. The packages are implemented in R with native C and
Fortran code. Using \c{cloc}, we measure a total \CranAllRCodeRnd lines of R
code and \CranAllNativeCodeRnd lines of native code. Out of these packages,
\CranPackages (\CranPackagesRatio) call eval. These packages represent
\CranCodeRatio of the code size.

To analyze a package, we need client code that invokes its methods. There are
three sources of built-in runnable code that come with each CRAN package:
\emph{tests, examples} and \emph{vignettes}. They are, respectively, traditional
unit tests, code snippets from the documentation, and long-form use-cases
written in \LaTeX\xspace or Markdown with executable snippets of R code.
Examples and vignettes are automatically extracted and turned into scripts,
their input is bundled with the package. Tests are directly runnable. Together
we there are \CranRunnableScripts scripts that contain \CranRunnableCode lines
of code (\CranRunnableCodeExamplesRnd in examples, \CranRunnableCodeVignettesRnd
in vignettes and \CranRunnableCodeTestsRnd in tests).

\mypara{Kaggle Scripts.} Kaggle is an online platform for data-science and
machine-learning. The website allows people to submit data-analysis problems,
users compete to find the best solution. The solutions, called \emph{kernels},
are uploaded to the platform as either plain scripts or notebooks. We obtained
\KaggleKernels kernels. Since Kaggle is not a curated repository, we first had
to remove duplicates. We used SHA-1 hash to identify \KaggleDuplicates
whole-file copies. The remaining \KaggleUnique kernels represent unique
solutions to \KaggleCompetitions Kaggle competition. Together, they contain
\KaggleCode lines of R code. Out of the \KaggleUniqueRnd kernels, only
\KaggleWithEvals contain calls to eval.

\subsection{Analysis Pipeline}

The results presented in this paper are the result of an automated analysis
pipeline that acquires the code of packages, extract metadata, executes
programs, traces their behavior and summarizes the observations.
Figure~\ref{fig:pipeline} shows the main steps of the pipeline along with
approximate time to execute each step, the data size, and the number of elements
manipulated by the stage. Timings are for runs on a cluster of three servers
with Intel Xeon 6140, 2.30GHz with 72 cores and 256GB of RAM using a shared OCFS
network storage.

\begin{figure}[!h]\hspace{-5mm}
  \includegraphics[width=1.05\linewidth]{pipeline.pdf}
  \caption{Analysis Pipeline UPDATE!!}\label{fig:pipeline}
\end{figure}

\noindent The pipeline has the following steps:
\begin{compactenum}
  \item \emph{Download.} The first step consists of downloading and installing
  CRAN packages and acquiring Kaggle corpus. We try to install all available
  CRAN packages. This is somewhat complicated by the fact that many of packages
  depend on native libraries, but these dependencies are not explicitly stated
  in the package description and thus it is hard to automate. For Kaggle we use
  a web crawler to get the programs and the a Kaggle official command line tool
  to download the datasets.
  \item \emph{Runnable code extraction.} Next step is to extract runnabale code.
  For CRAN package we use \genthat~\cite{issta18}, a tool which converts
  runnable code snippets from package documentation and vignettes as well as R
  tests into standalone executable scripts. For Kaggle programs that are not
  already R scripts, we use \c{rmarkdown} and \c{knitr} R packages to extract
  the runnable code from the Rmarkdown or the Jupyter notebook formats
  respectively. Duplicated Kaggle programs are removed. Each of the extracted
  script is wrapped into a call to our dynamic analyzer so we only record eval
  usage triggered by running th script. Without this, the data would include
  eval calls from the unit testing frameworks as well as from bootstrapping R
  virtual machine itself.
  \item \emph{Tracing.} With the scripts ready, we run our dynamic analysis that
  traces the calls to eval and captures other runtime information. For the CRAN
  packages we run two configurations:
  \begin{inparaenum}[(1)]
    \item all of the \CranRunnableScriptsRnd scripts while tracing calls to eval
    originating in any of the \CranAllPackages installed CRAN package, and
    \item a random selection of 1/4 of the scripts while tracing only eval calls
    originating in the core R packages that come bundled with the language.
  \end{inparaenum}
  The reason for this split is that the core R packages are intertwined with the
  language implementation and the usage is different form the user written
  packages hosted on CRAN. For Kaggle, we only trace the eval calls that
  originate in the programs themselves. To avoid any interference, each program
  is run in isolation, in its own R instance.
  \item \emph{Analysis.} In the final step, the output from the dynamic analysis
  is merged, cleaned and summarized in the post-processing phase by custom R
  scripts. The summarized data is then analyzed in RMarkdown notebooks to gather
  insights. Apart from the figures, the data points included in the paper are
  also generated by RMarkdown notebooks as \LaTeX macros.
\end{compactenum}

The code extraction and tracing steps of the pipeline are run in parallel on
three servers using GNU parallel~\cite{GNUparallel}. To make sure, all three
servers have identical environment, we use docker images that have all the
dependencies installed. The pipeline is implemented in 500 lines of Makefile and
3K lines of auxiliary R and shell scripts. Their main function is to gracefully
handle and report failures across large-scale program runs which greatly aids
debugging of the pipeline.

\subsection{Dynamic Analysis}

@@

\mypara{Limitations.}
%
% We turn off the bytecode compiler for this study. The bytecode
% compiler can also call \eval.
%
We do not get source locations for \CranUndefinedRnd \eval calls. In
these cases \eval is either passed as an argument to a higher-order functions or
is defined in a function returned by a higher-order function, or it can be a
generated called from package native code. However, this is a meager
\CranUndefinedRatio of all \eval calls and is unlikely to affect our
analysis. We ignore calls to the native \eval function exposed by R. We also
ignore the \c{rlang::tidy\_eval} function which uses native \eval internally
because \c{rlang} is used to implement a DSL for data analysis in R. It
introduces a new first-class promise object called \c{quosure} for which it
implements special evaluation support in \c{tidy\_eval}.

>>>>

This section explains how we selected our corpus and how we obtained the
reported results.

@@

Our corpus is assembled out of two set of programs, one consisting of libraries
obtained from \emph{CRAN} (Comprehensive R Archive Network\footnote{CRAN is the
  largest repository of R code with almost \CranAvailablePackagesRnd packages.
  It receives about 6 new package submissions a day~\cite{Ligges2017}. \cf
  \url{https://cran.r-project.org}}) and the other from
the \emph{Kaggle} data science competition website\footnote{\cf
  \url{https://kaggle.com}}. The intent here is to contrast code written by
experienced developers (CRAN) with code authored by typical end users of the
language (Kaggle). Another difference is that CRAN is a curated repository. Each
submitted package must abide to a number of well-formedness rules that are
automatically checked asserting certain quality. Most relevant for this work is
that all of the runnable code is tested and only a successfully running package
is admitted in the archive. Kaggle on the other hand is more like GitHub, anyone
can submit any code. All the code in our corpus is runnable, \ie we have scripts
that invoke the CRAN packages and the Kaggle programs have input data.

\mypara{Packages.} We obtained \CranAllPackages packages from CRAN, only
excluding packages that did not install due to missing native dependencies.
Using \c{cloc}, we measured \CranAllRCodeRnd lines of R and
\CranAllNativeCodeRnd lines of native code. Out of these packages, \CranPackages
(\CranPackagesRatio) call eval. These packages represent \CranCodeRatio of the
code size. To analyze a package, we need client code that invokes its methods.
There are three sources of built-in runnable code that come with each CRAN
package: \emph{tests, examples} and \emph{vignettes}. They are, respectively,
traditional unit tests, code snippets from the documentation, and long-form
use-cases written in \LaTeX\xspace or Markdown with executable snippets of R
code. Examples and vignettes are automatically extracted and turned into
scripts, their input is bundled with the package. Tests are directly runnable.
Together we there are \CranRunnableScripts scripts that contain
\CranRunnableCode lines of code (\CranRunnableCodeExamplesRnd in examples,
\CranRunnableCodeVignettesRnd in vignettes and \CranRunnableCodeTestsRnd in
tests).

\mypara{Scripts.} Kaggle is an online platform for data science. It allows users
to submit problems and compete to solve them. Solutions, called \emph{kernels},
are uploaded as plain scripts or notebooks. We obtained \KaggleKernels kernels.
Since Kaggle is not curated, used SHA-1 hashes to identify and remove
\KaggleDuplicates duplicate entries. The remaining \KaggleUnique kernels are
unique solutions to \KaggleCompetitions competition. They contain \KaggleCode
lines of R code. Only \KaggleWithEvals kernels  call \eval.

\subsection{Pipeline}

The analysis presented in this has been automated by a pipeline that acquires
package code, extract metadata, executes programs, traces their behavior and
summarizes observations. Figure~\ref{fig:pipeline} shows the main steps of
analysis along with their running time, data size, and number of elements
manipulated. Timings are from a cluster of three 2.3GHz Intel Xeon 6140 servers,
with 72 cores and 256GB of RAM, and shared OCFS network storage.

\begin{figure}[!h]\hspace{-5mm}
  \includegraphics[width=1.05\linewidth]{pipeline.pdf}
  \caption{Pipeline UPDATE!!}\label{fig:pipeline}
\end{figure}

\noindent The pipeline steps are:
\begin{compactenum}
  \item \emph{Download.} The first step consists of downloading and installing
    CRAN packages and acquiring Kaggle corpus. We try to install all available
    CRAN packages. This is somewhat complicated by the fact that many of
    packages depend on native libraries, but these dependencies are not
    explicitly stated in the package description and thus it is hard to
    automate. For Kaggle we use a web crawler to get the programs and the a
    Kaggle official command line tool to download the datasets.
  \item \emph{Runnable code extraction.} Next step is to extract runnabale code.
    For CRAN package we use \genthat~\cite{issta18}, a tool which converts
    runnable code snippets from package documentation and vignettes as well as R
    tests into standalone executable scripts. For Kaggle programs that are not
    already R scripts, we use \c{rmarkdown} and \c{knitr} R packages to extract
    the runnable code from the Rmarkdown or the Jupyter notebook formats
    respectively. Duplicated Kaggle programs are removed. Each of the extracted
    script is wrapped into a call to our dynamic analyzer so we only record eval
    usage triggered by running th script. Without this, the data would include
    eval calls from the unit testing frameworks as well as from bootstrapping R
    virtual machine itself.
  \item \emph{Tracing.} With the scripts ready, we run our dynamic analysis that
    traces the calls to eval and captures other runtime information. For the
    CRAN packages we run two configurations:
  \begin{inparaenum}[(1)]
    \item all of the \CranRunnableScriptsRnd scripts while tracing calls to eval
      originating in any of the \CranAllPackages installed CRAN package, and
    \item a random selection of 1/4 of the scripts while tracing only eval calls
      originating in the core R packages that come bundled with the language.
  \end{inparaenum}
    The reason for this split is that the core R packages are intertwined with
    the language implementation and the usage is different form the user written
    packages hosted on CRAN. For Kaggle, we only trace the eval calls that
    originate in the programs themselves. To avoid any interference, each
    program is run in isolation, in its own R instance.
  \item \emph{Analysis.} In the final step, the output from the dynamic analysis
    is merged, cleaned and summarized in the post-processing phase by custom R
    scripts. The summarized data is then analyzed in RMarkdown notebooks to
    gather insights. Apart from the figures, the data points included in the
    paper are also generated by RMarkdown notebooks as \LaTeX macros.
\end{compactenum}

\medskip\noindent
The code extraction and tracing steps of the pipeline are run in
parallel~\cite{GNUparallel}. Servers have identical environment thanks to docker
images with all dependencies installed. The pipeline consists of 3.5K lines of
Make, R and shell scripts.

\subsection{Analysis}

@@

\mypara{Limitations.} We do not get source locations for \CranUndefinedRnd \eval
calls. In these cases \eval is either passed as an argument to a higher-order
functions or is defined in a function returned by a higher-order function, or it
can be a generated called from package native code. However, this is a meager
\CranUndefinedRatio of all \eval calls and is unlikely to affect our analysis.
We ignore calls to the native \eval function exposed by R. We also ignore the
\c{rlang::tidy\_eval} function which uses native \eval internally because
\c{rlang} is used to implement a DSL for data analysis in R. It introduces a new
first-class promise object called \c{quosure} for which it implements special
evaluation support in \c{tidy\_eval}.
